---
title: "BFAST0N comparison vs BFAST and BFASTMonitor"
author: "Dainius Masiliūnas"
date: "22/02/2021"
output: html_document
---

```{r setup, include=FALSE}
library(DT)
library(future)
plan("multicore")
source("../src/bfast-cal/01-preprocess.r")
source("../src/bfast-cal/02-detectbreaks.r")
source("../src/bfast-cal/03-batchprocessing.r")
source("../src/bfast-cal/perclass.r")
source("../src/bfast-cal/utils.r")
source("../src/utils/visualise_breakpoints.r")
source("../src/utils/enable_fast_bfast.r")
```
```{r load, cache=TRUE}
GlobalData = LoadReferenceData("../data/Data_Global.csv")
```
```{r extractdata}
GlobalTS = LoadVITS(GlobalData, "", "/data/cgl_lc_bfast/Modis_VIs_GLOBAL_UTM_FINAL/", "Global")
```
```{r mergedata, cache=TRUE}
GlobalData = MergeAllYears(GlobalData, GlobalTS)
CheckReferenceData(GlobalData)
```
```{r addchangetypes}
GlobalData$continent = NULL
names(GlobalData)[names(GlobalData)=="field_30"] = "continent"
table(GlobalData$continent)
GlobalData[GlobalData$dominant_lc=="wetland_hebaceous",]$dominant_lc = "wetland_herbaceous"
GlobalData = AddChangeClassCol(GlobalData)
GlobalData = AddChangeProcessCol(GlobalData)
GlobalData = GlobalData[GlobalData$reference_year %in% 2016:2018,] # Discard any changes in 2015, reference does not track that
```

```{r selectchange}
ChangeData = GlobalData[GlobalData$sample_id %in% GlobalData[GlobalData$change_at_300m == "yes",]$sample_id,]
```

```{r}
table(GlobalData$changeprocess)
sort(table(GlobalData$changeclass), decreasing = TRUE)
```

# Overall statistics

Number of unique locations:
```{r}
length(unique(GlobalData$sample_id))
length(unique(GlobalData$location_id))
nrow(GlobalData) / 3
```

Number of instances of change:

```{r}
sum(GlobalData$change_at_300m == "yes")
sum(!is.na(GlobalData$changeclass))
```

Number of unique locations with change:

```{r}
length(unique(GlobalData[GlobalData$change_at_300m == "yes",]$sample_id))
```

Number of observations at locations with change:

```{r}
nrow(ChangeData)
```

Change to no change ratio:

```{r}
sum(ChangeData$change_at_300m == "yes") / nrow(ChangeData)
```

# Narrowing points down further

Test all points with a strucchange test and discard those with no break in the time series altogether (NB: the break can be outside our scope of 2016-2018!)

Example for one point:

```{r getonebreak}
SampleTS = sample(1:nrow(ChangeData), 1)
PlotTS(ChangeData[SampleTS,])
IsThereABreak(ChangeData[SampleTS,])
IsThereABreak(ChangeData[SampleTS,], formula=response~trend)
IsThereABreak(ChangeData[SampleTS,], formula=response~harmon)
```

Run for all points and see how many remain (using defaults):

```{r breakexists, cache=TRUE}
suppressMessages({BreakInBoth = apply(ChangeData[!duplicated(ChangeData$sample_id),], 1, IsThereABreak)})
suppressMessages({BreakInTrend = apply(ChangeData[!duplicated(ChangeData$sample_id),], 1, IsThereABreak, formula=response~trend)})
suppressMessages({BreakInSeasonality = apply(ChangeData[!duplicated(ChangeData$sample_id),], 1, IsThereABreak, formula=response~harmon)})
BreakInBoth[is.na(BreakInBoth)] = FALSE
BreakInTrend[is.na(BreakInTrend)] = FALSE
BreakInSeasonality[is.na(BreakInSeasonality)] = FALSE
BreakInAll = BreakInBoth & BreakInTrend & BreakInSeasonality
sum(BreakInAll, na.rm = TRUE)
```

How accurate is the sctest? Let's see a few examples:

```{r}
BreakIDs = unique(ChangeData$sample_id[BreakInAll])
BreakExamples = sample(BreakIDs, 20)
for (i in 1:20)
    PlotTS(ChangeData[ChangeData$sample_id == BreakExamples[i],])
```

Save the smaller dataset to file so that we can go back to it later.

```{r}
BreakData = ChangeData[ChangeData$sample_id %in% BreakIDs,]
st_write(BreakData, "../data/bfast0n-comparison-points-likebfast.gpkg", delete_dsn = TRUE)
```

## Test with filtering based on targetted sctest with both compotents and high confidence

```{r targetbreak, cache=TRUE}
suppressMessages({TargettedBreakPvals = apply(ChangeData[!duplicated(ChangeData$sample_id),], 1, MySctest, scsig=0.001, scrange=c(2015, 2020))})
TargettedBreakDF = data.frame(sample_id = unique(ChangeData$sample_id), pval = TargettedBreakPvals)

TargettedBreakDF = TargettedBreakDF[order(TargettedBreakDF$pval),]

BreakIDs = head(TargettedBreakDF, 1000)
print(tail(BreakIDs, 1))
BreakIDs = BreakIDs$sample_id
```

```{r}
BreakExamples = sample(BreakIDs, 20)
for (i in 1:20)
    PlotTS(ChangeData[ChangeData$sample_id == BreakExamples[i],])
```

```{r}
TargetData = ChangeData[ChangeData$sample_id %in% BreakIDs,]
st_write(TargetData, "../data/bfast0n-comparison-points-targetted.gpkg", delete_dsn = TRUE)
```

Select also 1000 no change points:

```{r targetnobreak, cache=TRUE}
NoChangeData = GlobalData[!GlobalData$sample_id %in% GlobalData[GlobalData$change_at_300m == "yes",]$sample_id,]
suppressMessages({TargettedNoBreakPvals = apply(NoChangeData[!duplicated(NoChangeData$sample_id),], 1, MySctest, scsig=0.001, scrange=c(2015, 2020))})
TargettedNoBreakDF = data.frame(sample_id = unique(NoChangeData$sample_id), pval = TargettedNoBreakPvals)

TargettedNoBreakDF = TargettedNoBreakDF[order(TargettedNoBreakDF$pval, decreasing=TRUE),]

# Remove points that are marked as changed in reference data
NoBreakIDs = head(TargettedNoBreakDF, 1000)
print(tail(NoBreakIDs, 1))
NoBreakIDs = NoBreakIDs$sample_id
```

```{r}
# Were any of the selected points marked as change?
any(NoChangeData[NoChangeData$sample_id %in% NoBreakIDs, ]$change_at_300m == "yes")
BreakExamples = sample(NoBreakIDs, 20)
for (i in 1:20)
    PlotTS(GlobalData[GlobalData$sample_id == BreakExamples[i],])
```
Save both sets:
```{r}
TargetChangeData = ChangeData[ChangeData$sample_id %in% BreakIDs,]
TargetNoChangeData = NoChangeData[NoChangeData$sample_id %in% NoBreakIDs, ]
TargetData = rbind(TargetChangeData, TargetNoChangeData)
st_write(TargetData, "../data/bfast0n-comparison-points-targetted-balanced.gpkg", delete_dsn = TRUE)
```

```{r}
rm(GlobalData, ChangeData, NoChangeData, BreakData, TargetChangeData, TargetNoChangeData, TargettedBreakDF, TargettedNoBreakDF, GlobalTS)
gc()
```

# Run change detection on each unique location

```{r, include=FALSE, eval=FALSE}
bf0nres.oa = TestMODBreakpointDetection(TargetData[(nrow(TargetData)-9):nrow(TargetData),], scrange=NULL, breaks="BIC", plot=TRUE, NewAccuracy = FALSE)
bfmres.oa = TestMODBreakpointDetection(TargetData[(nrow(TargetData)-9):nrow(TargetData),], DetectFunction = TestMODMonitor, plot=TRUE, NewAccuracy = FALSE)
```

```{r bf0nres, cache=TRUE, warning=FALSE, error=FALSE, include=FALSE, eval=FALSE}
bf0nres = TestMODBreakpointDetection(BreakData, scrange=NULL, breaks="BIC", NewAccuracy = TRUE, threshold = 0.5, order=3, quiet = TRUE)
```
```{r bfmres, cache=TRUE, include=FALSE, eval=FALSE}
bfmres = TestMODBreakpointDetection(BreakData, DetectFunction = TestMODMonitor, NewAccuracy = TRUE, threshold = 0.5, order=3, quiet=TRUE)
```
```{r bfres, cache=TRUE, warning=FALSE, error=FALSE, message=FALSE, results="hide", include=FALSE, eval=FALSE}
bfres = TestMODBreakpointDetection(BreakData, DetectFunction = TestMODBFAST, NewAccuracy = TRUE, threshold = 0.5, order=3, season="harmonic", quiet = TRUE)
```

```{r bfresInterp, cache=TRUE, warning=FALSE, error=FALSE, message=FALSE, results="hide", include=FALSE, eval=FALSE}
bfresInterp = TestMODBreakpointDetection(BreakData, DetectFunction = TestMODBFAST, NewAccuracy = TRUE, threshold = 0.5, order=3, season="harmonic", stlplus=FALSE, quiet=TRUE)
```

One iteration:
```{r bfres1, cache=TRUE, warning=FALSE, error=FALSE, message=FALSE, results="hide", include=FALSE, eval=FALSE}
bfres1 = TestMODBreakpointDetection(BreakData, DetectFunction = TestMODBFAST, NewAccuracy = TRUE, threshold = 0.5, order=3, season="harmonic", quiet = TRUE, max.iter=1)
```

```{r bfresInterp1, cache=TRUE, warning=FALSE, error=FALSE, message=FALSE, results="hide", include=FALSE, eval=FALSE}
bfresInterp1 = TestMODBreakpointDetection(BreakData, DetectFunction = TestMODBFAST, NewAccuracy = TRUE, threshold = 0.5, order=3, season="harmonic", stlplus=FALSE, quiet=TRUE, max.iter=1)
```

## Test on targetted and balanced breaks

Low tolerance to see whether bfastMonitor is really that good

```{r bf0ntarget, cache=TRUE, warning=FALSE, error=FALSE}
bf0ntarget = TestMODBreakpointDetection(TargetData, scrange=NULL, breaks="BIC", NewAccuracy = TRUE, threshold = 0.25, order=3, quiet = TRUE)
```
```{r bfmtarget, cache=TRUE}
bfmtarget = TestMODBreakpointDetection(TargetData, DetectFunction = TestMODMonitor, NewAccuracy = TRUE, threshold = 0.25, order=3, quiet=TRUE)
```
```{r bftarget, cache=TRUE, warning=FALSE, error=FALSE, message=FALSE, results="hide"}
bftarget = TestMODBreakpointDetection(TargetData, DetectFunction = TestMODBFAST, NewAccuracy = TRUE, threshold = 0.25, order=3, season="harmonic", quiet = TRUE)

```

```{r bftargetInterp, cache=TRUE, warning=FALSE, error=FALSE, message=FALSE, results="hide"}
bftargetInterp = TestMODBreakpointDetection(TargetData, DetectFunction = TestMODBFAST, NewAccuracy = TRUE, threshold = 0.25, order=3, season="harmonic", stlplus=FALSE, quiet=TRUE)
```

One iteration didn't do much difference
```{r bftarget1, cache=TRUE, warning=FALSE, error=FALSE, message=FALSE, results="hide", include=FALSE, eval=FALSE}
bftarget1 = TestMODBreakpointDetection(TargetData, DetectFunction = TestMODBFAST, NewAccuracy = TRUE, threshold = 0.25, order=3, season="harmonic", quiet = TRUE, max.iter=1)
```

```{r bftargetInterp1, cache=TRUE, warning=FALSE, error=FALSE, message=FALSE, results="hide", include=FALSE, eval=FALSE}
bftargetInterp1 = TestMODBreakpointDetection(TargetData, DetectFunction = TestMODBFAST, NewAccuracy = TRUE, threshold = 0.25, order=3, season="harmonic", stlplus=FALSE, quiet=TRUE, max.iter=1)
```

## Retry with tolerance=1

It seems that running with max.iter=1 is not very useful, and which points we test on doesn't make a huge difference either, so stick with base

```{r bf0nres1y, cache=TRUE, warning=FALSE, error=FALSE}
bf0nres1y = TestMODBreakpointDetection(TargetData, scrange=NULL, breaks="BIC", NewAccuracy = TRUE, threshold = 1, order=3, quiet = TRUE)
```
```{r bfmres1y, cache=TRUE}
bfmres1y = TestMODBreakpointDetection(TargetData, DetectFunction = TestMODMonitor, NewAccuracy = TRUE, threshold = 1, order=3, quiet=TRUE)
```
```{r bfres1y, cache=TRUE, warning=FALSE, error=FALSE, message=FALSE, results="hide"}
bfres1y = TestMODBreakpointDetection(TargetData, DetectFunction = TestMODBFAST, NewAccuracy = TRUE, threshold = 1, order=3, season="harmonic", quiet = TRUE)

```

```{r bfresInterp1y, cache=TRUE, warning=FALSE, error=FALSE}
bfresInterp1y = TestMODBreakpointDetection(TargetData, DetectFunction = TestMODBFAST, NewAccuracy = TRUE, threshold = 1, order=3, season="harmonic", stlplus=FALSE, quiet=TRUE)
```

BFAST Lite with LWZ:

```{r bf0nres1ylwz, cache=TRUE, warning=FALSE, error=FALSE}
bf0nres1ylwz = TestMODBreakpointDetection(TargetData, scrange=NULL, breaks="LWZ", NewAccuracy = TRUE, threshold = 1, order=3, quiet = TRUE)
```

## Rerun with yearly statistics

The results are surprising; is it because we are missing the middle, or is it because of the new accuracy method? Try again with the old method.

```{r bf0nres1yoa, cache=TRUE, warning=FALSE, error=FALSE}
bf0nres1y.oa = TestMODBreakpointDetection(TargetData, scrange=NULL, breaks="BIC", NewAccuracy = FALSE, threshold = 1, order=3, quiet = TRUE)
```
```{r bfmres1yoa, cache=TRUE}
bfmres1y.oa = TestMODBreakpointDetection(TargetData, DetectFunction = TestMODMonitor, NewAccuracy = FALSE, threshold = 1, order=3, quiet=TRUE)
```
```{r bfres1yoa, cache=TRUE, warning=FALSE, error=FALSE, message=FALSE, results="hide"}
bfres1y.oa = TestMODBreakpointDetection(TargetData, DetectFunction = TestMODBFAST, NewAccuracy = FALSE, threshold = 1, order=3, season="harmonic", quiet = TRUE)
```

```{r bfresInterp1yoa, cache=TRUE, warning=FALSE, error=FALSE}
bfresInterp1y.oa = TestMODBreakpointDetection(TargetData, DetectFunction = TestMODBFAST, NewAccuracy = FALSE, threshold = 1, order=3, season="harmonic", stlplus=FALSE, quiet=TRUE)
```

```{r bf0nres1ylwzoa, cache=TRUE, warning=FALSE, error=FALSE}
bf0nres1ylwz.oa = TestMODBreakpointDetection(TargetData, scrange=NULL, breaks="LWZ", NewAccuracy = FALSE, threshold = 1, order=3, quiet = TRUE)
```


## Accuracy comparison

```{r}
datatable(rbind(
    #bf0n = FPStats(bf0nres),
    #bfm = FPStats(bfmres),
    #bfstlp=FPStats(bfres),
    #bfinterp=FPStats(bfresInterp),
    #bfstlp1=FPStats(bfres1),
    #bfinterp1=FPStats(bfresInterp1),
    bf0n.025y = FPStats(bf0ntarget),
    bfm.025y = FPStats(bfmtarget),
    bfstlp.025y=FPStats(bftarget),
    bfinterp.025y=FPStats(bftargetInterp),
    #bfstlp1.t=FPStats(bftarget1),
    #bfinterp1.t=FPStats(bftargetInterp1),
    bf0n.1y = FPStats(bf0nres1y),
    bfm.1y = FPStats(bfmres1y),
    bfstlp.1y=FPStats(bfres1y),
    bfinterp.1y=FPStats(bfresInterp1y),
    bf0n.1y.lwz=FPStats(bf0nres1ylwz),
    bf0n.1y.oa = FPStats(bf0nres1y.oa),
    bfm.1y.oa = FPStats(bfmres1y.oa),
    bfstlp.1y.oa=FPStats(bfres1y.oa),
    bfinterp.1y.oa=FPStats(bfresInterp1y.oa),
    bf0n.1y.lwz.oa=FPStats(bf0nres1ylwz.oa)
    ))
```
