---
title: "BFAST0N comparison vs BFAST and BFASTMonitor"
author: "Dainius Masiliūnas"
date: "22/02/2021"
output: html_document
---

```{r setup, include=FALSE}
library(DT)
library(pbapply)
source("../src/bfast-cal/01-preprocess.r")
source("../src/bfast-cal/02-detectbreaks.r")
source("../src/bfast-cal/03-batchprocessing.r")
source("../src/bfast-cal/perclass.r")
source("../src/bfast-cal/utils.r")
source("../src/utils/visualise_breakpoints.r")
```
```{r load, cache=TRUE}
GlobalData = LoadReferenceData("../data/Data_Global.csv")
```
```{r extractdata}
GlobalTS = LoadVITS(GlobalData, "", "/data/cgl_lc_bfast/Modis_VIs_GLOBAL_UTM_FINAL/", "Global")
```
```{r mergedata, cache=TRUE}
GlobalData = MergeAllYears(GlobalData, GlobalTS)
CheckReferenceData(GlobalData)
```
```{r addchangetypes}
GlobalData$continent = NULL
names(GlobalData)[names(GlobalData)=="field_30"] = "continent"
table(GlobalData$continent)
GlobalData[GlobalData$dominant_lc=="wetland_hebaceous",]$dominant_lc = "wetland_herbaceous"
GlobalData = AddChangeClassCol(GlobalData)
GlobalData = AddChangeProcessCol(GlobalData)
GlobalData = GlobalData[GlobalData$reference_year %in% 2016:2018,] # Discard any changes in 2015, reference does not track that
```

```{r selectchange}
ChangeData = GlobalData[GlobalData$sample_id %in% GlobalData[GlobalData$change_at_300m == "yes",]$sample_id,]
```

```{r}
table(GlobalData$changeprocess)
sort(table(GlobalData$changeclass), decreasing = TRUE)
```

# Overall statistics

Number of unique locations:
```{r}
length(unique(GlobalData$sample_id))
length(unique(GlobalData$location_id))
nrow(GlobalData) / 3
```

Number of instances of change:

```{r}
sum(GlobalData$change_at_300m == "yes")
sum(!is.na(GlobalData$changeclass))
```

Number of unique locations with change:

```{r}
length(unique(GlobalData[GlobalData$change_at_300m == "yes",]$sample_id))
```

Number of observations at locations with change:

```{r}
nrow(ChangeData)
```

Change to no change ratio:

```{r}
sum(ChangeData$change_at_300m == "yes") / nrow(ChangeData)
```

# Narrowing points down further

Test all points with a strucchange test and discard those with no break in the time series altogether (NB: the break can be outside our scope of 2016-2018!)

Example for one point:

```{r getonebreak}
PlotTS(ChangeData[1,])
IsThereABreak(ChangeData[1,])
```

Run for all points and see how many remain (using defaults):

```{r breakexists, cache=TRUE}
suppressMessages({BreakExists = apply(ChangeData[!duplicated(ChangeData$sample_id),], 1, IsThereABreak)})
sum(BreakExists, na.rm = TRUE)
```

How accurate is the sctest? Let's see a few examples:

```{r}
BreakExists[is.na(BreakExists)] = FALSE
BreakIDs = unique(ChangeData$sample_id)[BreakExists]
BreakExamples = sample(BreakIDs, 20)
for (i in 1:20)
    PlotTS(ChangeData[ChangeData$sample_id == BreakExamples[i],])
```

It is a bit lenient. We can survive even with ~10000 unique locations. Try harder.

```{r}
suppressMessages({BreakExists = apply(ChangeData[!duplicated(ChangeData$sample_id),], 1, IsThereABreak, scsig=0.001, scrange=c(2015, 2020))})
sum(BreakExists, na.rm = TRUE)
BreakExists[is.na(BreakExists)] = FALSE
BreakIDs = unique(ChangeData$sample_id)[BreakExists]
BreakExamples = sample(BreakIDs, 20)
for (i in 1:20)
    PlotTS(ChangeData[ChangeData$sample_id == BreakExamples[i],])
```

Save the smaller dataset to file so that we can go back to it later.

```{r}
BreakData = ChangeData[ChangeData$sample_id %in% BreakIDs,]
st_write(BreakData, "../data/bfast0n-comparison-points.gpkg", delete_dsn = TRUE)
rm(GlobalData, GlobalTS, ChangeData)
```

# Run BFAST0N on each unique location

```{r}
bf0nres = TestMODBreakpointDetection(BreakData[1:9,], scrange=NULL, breaks="BIC", plot=TRUE)
bfmres = TestMODBreakpointDetection(BreakData[1:9,], DetectFunction = TestMODMonitor)
```

New way to get accuracy statistics:

```{r}
bf0nres = TestMODBreakpointDetection(BreakData[1:9,], scrange=NULL, breaks="BIC", plot=TRUE, NewAccuracy = TRUE, threshold = 0.5)
bfmres = TestMODBreakpointDetection(BreakData[1:9,], DetectFunction = TestMODMonitor, NewAccuracy = TRUE, threshold = 0.5)
```
