---
title: "Forest change optimisation"
author: "Dainius"
date: '2020-03-18'
output: html_document
---

```{r setup, include=FALSE}
library(DT)
source("../src/bfast-cal/01-preprocess.r")
source("../src/bfast-cal/02-detectbreaks.r")
source("../src/bfast-cal/perclass.r")
source("../src/bfast-cal/utils.r")
```
```{r load, cache=TRUE}
GlobalData = LoadReferenceData("../data/NIRv_16d_byte_full_Global.csv")
```
```{r addchange, cache=TRUE}
GlobalData = AddChangeClassCol(GlobalData)
```

```{r table}
sort(table(GlobalData$changeclass), TRUE)
```

```{r trees_subset}
TreesChange = FilterChange(GlobalData, "tree to grassland")
#Subset to only one continent for now
#TreesChange = TreesChange[TreesChange$continent=="North America",]
```

```{r samples, eval=FALSE}
table(TreesChange[TreesChange$sample_id %in% TreesChange[TreesChange$changeclass %in% "tree to grassland",]$sample_id,]$sample_id)
```

First example: missing values in certain years
```{r plot_bfast}
SampleTS = GetMatrixFromSF(TreesChange[TreesChange$sample_id==1360691,])[1,]
MODDetectBreaks(SampleTS, order=3, scrange=NULL, plot=TRUE, TargetYears=TreesChange[TreesChange$sample_id==1360691 & TreesChange$change_at_300m=="yes", ]$reference_year)
```
Also a lot of missing values

```{r plot_bfast2}
Sample = 1360997
SampleTS = GetMatrixFromSF(TreesChange[TreesChange$sample_id==Sample,])[1,]
MODDetectBreaks(SampleTS, order=3, scrange=NULL, plot=TRUE, TargetYears=TreesChange[TreesChange$sample_id==Sample & TreesChange$change_at_300m=="yes", ]$reference_year)
```

```{r plot_bfast3}
Sample = 1363062
SampleTS = GetMatrixFromSF(TreesChange[TreesChange$sample_id==Sample,])[1,]
MODDetectBreaks(SampleTS, order=3, scrange=NULL, plot=TRUE, TargetYears=TreesChange[TreesChange$sample_id==Sample & TreesChange$change_at_300m=="yes", ]$reference_year)
```

```{r plot_bfast4}
Sample = 1114232
SampleTS = GetMatrixFromSF(TreesChange[TreesChange$sample_id==Sample,])[1,]
MODDetectBreaks(SampleTS, order=3, scrange=NULL, plot=TRUE, TargetYears=TreesChange[TreesChange$sample_id==Sample & TreesChange$change_at_300m=="yes", ]$reference_year)
```
In Asia
```{r plot_bfast5}
Sample = 1372873
SampleTS = GetMatrixFromSF(GlobalData[GlobalData$sample_id==Sample,])[1,]
MODDetectBreaks(SampleTS, order=3, scrange=NULL, plot=TRUE, TargetYears=GlobalData[GlobalData$sample_id==Sample & GlobalData$change_at_300m=="yes", ]$reference_year)
```
In South America

```{r plot_bfast6}
Sample = 1390578
SampleTS = GetMatrixFromSF(GlobalData[GlobalData$sample_id==Sample,])[1,]
MODDetectBreaks(SampleTS, order=3, scrange=NULL, plot=TRUE, TargetYears=GlobalData[GlobalData$sample_id==Sample & GlobalData$change_at_300m=="yes", ]$reference_year)
```

In North America

```{r plot_bfast7}
Sample = 1376056
SampleTS = GetMatrixFromSF(GlobalData[GlobalData$sample_id==Sample,])[1,]
MODDetectBreaks(SampleTS, order=3, scrange=NULL, plot=TRUE, TargetYears=GlobalData[GlobalData$sample_id==Sample & GlobalData$change_at_300m=="yes", ]$reference_year)
```

Check whether we have any complete cases at all
```{r}
ChangeOnly = GlobalData[GlobalData$change_at_300m=="yes",]
ChangeOnly[which(!is.na(ChangeOnly$X2015.01.03)),]$sample_id
```

That is sad (also very weird that these samples are OK)

```{r plot_bfast8}
Sample = 1399440
SampleTS = GetMatrixFromSF(GlobalData[GlobalData$sample_id==Sample,])[1,]
MODDetectBreaks(SampleTS, order=3, scrange=NULL, plot=TRUE, TargetYears=GlobalData[GlobalData$sample_id==Sample & GlobalData$change_at_300m=="yes", ]$reference_year)
```

```{r plot_bfast10}
Sample = 1154715
SampleTS = GetMatrixFromSF(GlobalData[GlobalData$sample_id==Sample,])[1,]
MODDetectBreaks(SampleTS, order=3, scrange=NULL, plot=TRUE, TargetYears=GlobalData[GlobalData$sample_id==Sample & GlobalData$change_at_300m=="yes", ]$reference_year)
```

```{r plot_bfast9}
Sample = 1121506
SampleTS = GetMatrixFromSF(GlobalData[GlobalData$sample_id==Sample,])[1,]
MODDetectBreaks(SampleTS, order=3, scrange=NULL, plot=TRUE, TargetYears=GlobalData[GlobalData$sample_id==Sample & GlobalData$change_at_300m=="yes", ]$reference_year)
```

Let's choose 1121506 for now. What if we change the order:

```{r plot_bfast_order2}
Sample = 1121506
SampleTS = GetMatrixFromSF(GlobalData[GlobalData$sample_id==Sample,])[1,]
MODDetectBreaks(SampleTS, order=2, scrange=NULL, plot=TRUE, TargetYears=GlobalData[GlobalData$sample_id==Sample & GlobalData$change_at_300m=="yes", ]$reference_year)
```

Order 1:
```{r plot_bfast_order1}
SampleTS = GetMatrixFromSF(GlobalData[GlobalData$sample_id==Sample,])[1,]
MODDetectBreaks(SampleTS, order=1, scrange=NULL, plot=TRUE, TargetYears=GlobalData[GlobalData$sample_id==Sample & GlobalData$change_at_300m=="yes", ]$reference_year)
```

No seasonality:
```{r plot_bfast_trendonly}
SampleTS = GetMatrixFromSF(GlobalData[GlobalData$sample_id==Sample,])[1,]
MODDetectBreaks(SampleTS, order=1, scrange=NULL, plot=TRUE, formula=response~trend, TargetYears=GlobalData[GlobalData$sample_id==Sample & GlobalData$change_at_300m=="yes", ]$reference_year)
```

Simple seasonality:
```{r plot_bfast_trendseason}
SampleTS = GetMatrixFromSF(GlobalData[GlobalData$sample_id==Sample,])[1,]
MODDetectBreaks(SampleTS, order=1, scrange=NULL, plot=TRUE, formula=response~trend+season, seasonfreq=0.5, TargetYears=GlobalData[GlobalData$sample_id==Sample & GlobalData$change_at_300m=="yes", ]$reference_year)
```

Equivalent on BIC:

```{r plot_bfast_bic}
SampleTS = GetMatrixFromSF(GlobalData[GlobalData$sample_id==Sample,])[1,]
MODDetectBreaks(SampleTS, order=3, scrange=NULL, plot=TRUE, breaks="BIC", TargetYears=GlobalData[GlobalData$sample_id==Sample & GlobalData$change_at_300m=="yes", ]$reference_year)
```

```{r plot_bfast_bic}
SampleTS = GetMatrixFromSF(GlobalData[GlobalData$sample_id==Sample,])[1,]
MODDetectBreaks(SampleTS, order=1, scrange=NULL, plot=TRUE, breaks="BIC", TargetYears=GlobalData[GlobalData$sample_id==Sample & GlobalData$change_at_300m=="yes", ]$reference_year)
```

```{r plot_bfast_trendonly_bic}
SampleTS = GetMatrixFromSF(GlobalData[GlobalData$sample_id==Sample,])[1,]
MODDetectBreaks(SampleTS, order=1, scrange=NULL, plot=TRUE, formula=response~trend, breaks="BIC", TargetYears=GlobalData[GlobalData$sample_id==Sample & GlobalData$change_at_300m=="yes", ]$reference_year)
```

```{r plot_bfast_trendseason_bic}
SampleTS = GetMatrixFromSF(GlobalData[GlobalData$sample_id==Sample,])[1,]
MODDetectBreaks(SampleTS, order=1, scrange=NULL, plot=TRUE, formula=response~trend+season, seasonfreq=0.5, breaks="BIC", TargetYears=GlobalData[GlobalData$sample_id==Sample & GlobalData$change_at_300m=="yes", ]$reference_year)
```

All in all, best model is trend only, no matter if 
