---
title: "CGLOPS global point testing"
author: "Dainius"
date: '2020-03-23'
output: html_document
---

```{r setup, include=FALSE}
library(DT)
source("../src/bfast-cal/01-preprocess.r")
source("../src/bfast-cal/02-detectbreaks.r")
source("../src/bfast-cal/03-batchprocessing.r")
source("../src/bfast-cal/perclass.r")
source("../src/bfast-cal/utils.r")
registerDoParallel(cores=4)
```
```{r load, cache=TRUE}
GlobalData = LoadReferenceData("../data/Data_Global.csv")
```
```{r extractdata}
GlobalTS = LoadVITS(GlobalData, "", "/data/cgl_lc_bfast/Modis_VIs_GLOBAL_UTM_FINAL/", "Global")
```
```{r, cache=TRUE}
GlobalData = MergeAllYears(GlobalData, GlobalTS)
CheckReferenceData(GlobalData)
```

```{r addchangetypes}
GlobalData$continent = NULL
names(GlobalData)[names(GlobalData)=="field_30"] = "continent"
table(GlobalData$continent)
GlobalData[GlobalData$dominant_lc=="wetland_hebaceous",]$dominant_lc = "wetland_herbaceous"
GlobalData = AddChangeClassCol(GlobalData)
GlobalData = AddChangeProcessCol(GlobalData)
GlobalData = GlobalData[GlobalData$reference_year %in% 2016:2018,] # Discard any changes in 2015, reference does not track that
```

```{r}
table(GlobalData$changeprocess)
sort(table(GlobalData$changeclass), decreasing = TRUE)
```

# Perform mass statistics

```{r baseline-stat-comparison, cache=TRUE}
ResultBaseline = TestParams(list(
    list(order=3, scrange=NULL),
    list(order=1, formula=response~trend, scrange=NULL),
    list(breaks="BIC", order=1, formula=response~trend),
    list(breaks="BIC", order=1, seasonfreq = 0.3, formula=response~trend+season, scrange=NULL)
    ), data=GlobalData)
```
## Overall results

```{r baseline-stats, cache=TRUE}
BS = FPStatsPerParam(ResultBaseline)
```

```{r}
datatable(BS)
```

## Results per class

```{r perclass-stats, cache=TRUE}
PCS = lapply(unique(ResultBaseline$call), function(x) FPStatsPerClass(ResultBaseline[ResultBaseline$call==x,], cl=2))
```
```{r}
htmltools::tagList(lapply(1:length(PCS), function(x)datatable(PCS[[x]], caption=unique(ResultBaseline$call)[x])))
rm(PCS)
```

The improvements are across the board, rather than in particular classes, with a few exceptions.

## Results per change process

```{r perprocess-stats, cache=TRUE}
PPS = lapply(unique(ResultBaseline$call), function(x) FPStatsPerClass(ResultBaseline[ResultBaseline$call==x,], column = "changeprocess", cl=2))
```
```{r}
htmltools::tagList(lapply(1:length(PPS), function(x)datatable(PPS[[x]], caption=unique(ResultBaseline$call)[x])))
rm(PPS)
```

As expected, between vegetation and non-vegetation it's much easier, and between vegetation is the hardest.

## Results per continent

```{r percontinent-stats}
htmltools::tagList(lapply(unique(ResultBaseline$call), function(x) datatable(FPStatsPerContinent(ResultBaseline[ResultBaseline$call==x,]), caption=x, cl=2)))
```

## Masking out unvegetated pixels

What if we don't run BFAST on unvegetated pixels?
```{r postprocess-vegetated}
ResultBaseline = AddVICol(ResultBaseline)
```

```{r vegetated15-stats}
ResultVegetated = ResultBaseline
ResultVegetated[ResultVegetated$nirv < 15,]$bfast_guess = FALSE
datatable(FPStatsPerParam(ResultVegetated))
```

What if we use a threshold of 0?

```{r vegetated0-stats}
ResultVegetated = ResultBaseline
ResultVegetated[ResultVegetated$nirv <= 0,]$bfast_guess = FALSE
datatable(FPStatsPerParam(ResultVegetated))
rm(ResultVegetated)
rm(ResultBaseline)
```

# Round 2: season frequency and sctest testing

```{r seasonality-comparison1, cache=TRUE}
system.time(ResultSeasonality1 <- TestMODBreakpointDetection(GlobalData, breaks="BIC", order=1, seasonfreq = 0.2, formula=response~trend+season, scrange=NULL))
```
Takes 16078s (4.5 hours)
```{r seasonality-comparison2, cache=TRUE}
system.time(ResultSeasonality2 <- TestMODBreakpointDetection(GlobalData, breaks="BIC", order=1, seasonfreq = 0.3, formula=response~trend+season, scrange=c(2009, 2020), scsig=0.01))
```
Takes 6886s (2 hours)
```{r seasonality-comparison3, cache=TRUE}
system.time(ResultSeasonality3 <- TestMODBreakpointDetection(GlobalData, breaks="BIC", order=1, seasonfreq = 0.3, formula=response~trend+season, scrange=c(2014, 2020), scsig=0.01))
```
Takes 5935s (1.6 hours)
```{r seasonality-comparison4, cache=TRUE}
system.time(ResultSeasonality4 <- TestMODBreakpointDetection(GlobalData, breaks="BIC", order=1, seasonfreq = 0.3, formula=response~trend+season, scrange=c(2014, 2020), scsig=0.005))
```
Takes 5309s (1.5 hours)

## Overall results

```{r}
AddExtraCols = function(data)
{
    data = AddVICol(data)
    data$changed = data$change_at_300m == "yes"
    data$comment = ""
    return(data)
}
ResultSeasonality1 = AddExtraCols(ResultSeasonality1)
ResultSeasonality1$call = "breaks BIC, seasonfreq = 0.2, formula=response~trend+season, scrange=NULL"
ResultSeasonality2 = AddExtraCols(ResultSeasonality2)
ResultSeasonality2$call = "breaks BIC, seasonfreq = 0.3, formula=response~trend+season, scrange=c(2009, 2020), scsig=0.01"
ResultSeasonality3 = AddExtraCols(ResultSeasonality3)
ResultSeasonality3$call = "breaks BIC, seasonfreq = 0.3, formula=response~trend+season, scrange=c(2014, 2020), scsig=0.01"
ResultSeasonality4 = AddExtraCols(ResultSeasonality4)
ResultSeasonality4$call = "breaks BIC, seasonfreq = 0.3, formula=response~trend+season, scrange=c(2014, 2020), scsig=0.005"
```

```{r seasonality-stats1, cache=TRUE}
RS1 = FPStatsPerParam(rbind(ResultSeasonality1, ResultSeasonality2, ResultSeasonality3, ResultSeasonality4))
```
```{r}
datatable(rbind(BS, RS1))
```

```{r include=FALSE, results='hide'}
rm(ResultSeasonality1, ResultSeasonality2, ResultSeasonality3, ResultSeasonality4)
gc()
```

# Round 3: additional sctest tests

Seasonality might also be possible to optimise, but focus on runtime for now because the results are very different.

```{r sctest-comparison1, cache=TRUE}
system.time(ResultSeasonality5 <- TestMODBreakpointDetection(GlobalData, quiet=TRUE, breaks="BIC", order=1, seasonfreq = 0.3, formula=response~trend+season, scrange=c(2009, 2020), scsig=0.1))
```
Took 11000s (3 hours)
```{r}
ResultSeasonality5 = AddExtraCols(ResultSeasonality5)
ResultSeasonality5$call = "breaks BIC, seasonfreq = 0.3, formula=response~trend+season, scrange=c(2009, 2020), scsig=0.1"
```
```{r sctest-comparison2, results='hide', cache=TRUE}
system.time(ResultSeasonality6 <- TestMODBreakpointDetection(GlobalData, quiet=TRUE, breaks="BIC", order=1, seasonfreq = 0.3, formula=response~trend+season, scrange=c(2014, 2020), scsig=0.1))
```
```{r}
ResultSeasonality6 = AddExtraCols(ResultSeasonality6)
ResultSeasonality6$call = "breaks BIC, seasonfreq = 0.3, formula=response~trend+season, scrange=c(2014, 2020), scsig=0.1"
```
```{r seasonality-stats2, cache=TRUE}
RS2 = FPStatsPerParam(rbind(ResultSeasonality5, ResultSeasonality6))
```
```{r include=FALSE, results='hide'}
rm(ResultSeasonality5, ResultSeasonality6)
gc()
```

```{r sctest-comparison3, cache=TRUE}
system.time(ResultSeasonality7 <- TestMODBreakpointDetection(GlobalData, quiet=TRUE, breaks="BIC", order=1, seasonfreq = 0.3, formula=response~trend+season, scrange=c(2009, 2020), scsig=0.3))
```
Took 14527s (4 hours)
```{r}
ResultSeasonality7 = AddExtraCols(ResultSeasonality7)
ResultSeasonality7$call = "breaks BIC, seasonfreq = 0.3, formula=response~trend+season, scrange=c(2009, 2020), scsig=0.3"
```
```{r sctest-comparison4, cache=TRUE}
system.time(ResultSeasonality8 <- TestMODBreakpointDetection(GlobalData, quiet=TRUE, breaks="BIC", order=1, seasonfreq = 0.3, formula=response~trend+season, scrange=c(2014, 2020), scsig=0.5))
```
Took 15671s (4.35 hours)
```{r}
ResultSeasonality8 = AddExtraCols(ResultSeasonality8)
ResultSeasonality8$call = "breaks BIC, seasonfreq = 0.3, formula=response~trend+season, scrange=c(2014, 2020), scsig=0.5"
```
```{r seasonality-stats3, cache=TRUE}
RS3 = FPStatsPerParam(rbind(ResultSeasonality7, ResultSeasonality8))
```

```{r}
datatable(rbind(BS, RS1, RS2, RS3))
```

```{r include=FALSE, results='hide'}
rm(ResultSeasonality7, ResultSeasonality8)
gc()
```
