---
title: "CGLOPS ROW Priority testing"
author: "Dainius"
date: '2020-03-20'
output: html_document
---

```{r setup, include=FALSE}
library(DT)
source("../src/bfast-cal/01-preprocess.r")
source("../src/bfast-cal/02-detectbreaks.r")
source("../src/bfast-cal/03-batchprocessing.r")
source("../src/bfast-cal/perclass.r")
source("../src/bfast-cal/utils.r")
registerDoParallel(cores=4)
```
```{r load, cache=TRUE}
GlobalData = LoadReferenceData("../data/Data_Global.csv")
```

```{r getrow}
ROWTS = LoadVITS(GlobalData, "", "/data/cgl_lc_bfast/Modis_VIs_ROWPR_UTM_FINAL/", "ROWPR")
ROWData = MergeAllYears(ROWTS, GlobalData)
ROWData = ROWData[order(ROWData$sample_id, ROWData$reference_year),]
CheckReferenceData(ROWData)
ROWData = AddChangeClassCol(ROWData)
```

```{r globalplot}
plot(st_geometry(ROWData))
```

```{r table}
sort(table(ROWData$changeclass), TRUE)
```

```{r trees_subset}
TreesChange = FilterChange(ROWData, "tree to grassland")
```

```{r samples}
ChangeSamples = unique(TreesChange[TreesChange$sample_id %in% TreesChange[TreesChange$changeclass %in% "tree to grassland",]$sample_id,]$sample_id)
length(ChangeSamples)
```

Let's look at 30 random points:

```{r plot_bfast1}
for (Sample in ChangeSamples[seq(1, length(ChangeSamples), 10)])
{
    print(Sample)
    SampleTS = GetMatrixFromSF(TreesChange[TreesChange$sample_id==Sample,])[1,]
    MODDetectBreaks(GetTS(SampleTS, frequency = 23), order=3, scrange=NULL, plot=TRUE, TargetYears=ROWData[ROWData$sample_id==Sample & ROWData$change_at_300m=="yes", ]$reference_year)
}
```

If we use order 2 LWZ:

```{r plot_bfast_order2_lwz}
for (Sample in ChangeSamples[seq(1, length(ChangeSamples), 10)])
{
    print(Sample)
    SampleTS = GetMatrixFromSF(TreesChange[TreesChange$sample_id==Sample,])[1,]
    MODDetectBreaks(GetTS(SampleTS, frequency = 23), order=2, scrange=NULL, plot=TRUE, TargetYears=ROWData[ROWData$sample_id==Sample & ROWData$change_at_300m=="yes", ]$reference_year)
}
```

This is clearly much better.

Try with trend only:

```{r plot_bfast_trend_lwz}
for (Sample in ChangeSamples[seq(1, length(ChangeSamples), 10)])
{
    print(Sample)
    SampleTS = GetMatrixFromSF(TreesChange[TreesChange$sample_id==Sample,])[1,]
    MODDetectBreaks(GetTS(SampleTS, frequency = 23), order=1, scrange=NULL, plot=TRUE, formula=response~trend, main=Sample, TargetYears=ROWData[ROWData$sample_id==Sample & ROWData$change_at_300m=="yes", ]$reference_year)
}
```

Also overall better (even than order 2), only in places with a lot of seasonality at the end of the time series it tends to estimate breaks too early.

Retry with BIC:

```{r plot_bfast_bic}
for (Sample in ChangeSamples[seq(1, length(ChangeSamples), 10)])
{
    print(Sample)
    SampleTS = GetMatrixFromSF(TreesChange[TreesChange$sample_id==Sample,])[1,]
    MODDetectBreaks(GetTS(SampleTS, frequency = 23), order=3, scrange=NULL, plot=TRUE, breaks="BIC", TargetYears=ROWData[ROWData$sample_id==Sample & ROWData$change_at_300m=="yes", ]$reference_year)
}
```

Too many breaks, and not as sensitive as trend, but also not too bad

BIC with order 2:

```{r plot_bfast_order2_bic}
for (Sample in ChangeSamples[seq(1, length(ChangeSamples), 10)])
{
    print(Sample)
    SampleTS = GetMatrixFromSF(TreesChange[TreesChange$sample_id==Sample,])[1,]
    MODDetectBreaks(GetTS(SampleTS, frequency = 23), order=2, scrange=NULL, plot=TRUE, breaks="BIC", formula=response~trend+harmon, main=Sample, TargetYears=ROWData[ROWData$sample_id==Sample & ROWData$change_at_300m=="yes", ]$reference_year)
}
```

Much better, may be better than trend LWZ

```{r plot_bfast_trend_bic}
for (Sample in ChangeSamples[seq(1, length(ChangeSamples), 10)])
{
    print(Sample)
    SampleTS = GetMatrixFromSF(TreesChange[TreesChange$sample_id==Sample,])[1,]
    MODDetectBreaks(GetTS(SampleTS, frequency = 23), order=1, scrange=NULL, plot=TRUE, breaks="BIC", formula=response~trend, main=Sample, TargetYears=ROWData[ROWData$sample_id==Sample & ROWData$change_at_300m=="yes", ]$reference_year)
}
```

Detects even the really tough ones, but breaks down when there is strong seasonality. Choose this for max sensitivity.

What if we use season to give very slight seasonality?

```{r plot_bfast_trendseason_lwz}
for (Sample in ChangeSamples[seq(1, length(ChangeSamples), 10)])
{
    print(Sample)
    SampleTS = GetMatrixFromSF(TreesChange[TreesChange$sample_id==Sample,])[1,]
    MODDetectBreaks(GetTS(SampleTS, frequency = 23), order=1, scrange=NULL, plot=TRUE, formula=response~trend+season, seasonfreq = 0.3, main=Sample, TargetYears=ROWData[ROWData$sample_id==Sample & ROWData$change_at_300m=="yes", ]$reference_year)
}
```

Decent, but not sensitive enough

```{r plot_bfast_trendseason_bic}
for (Sample in ChangeSamples[seq(1, length(ChangeSamples), 10)])
{
    print(Sample)
    SampleTS = GetMatrixFromSF(TreesChange[TreesChange$sample_id==Sample,])[1,]
    MODDetectBreaks(GetTS(SampleTS, frequency = 23), order=1, scrange=NULL, plot=TRUE, formula=response~trend+season, seasonfreq = 0.3, breaks = "BIC", main=Sample, TargetYears=ROWData[ROWData$sample_id==Sample & ROWData$change_at_300m=="yes", ]$reference_year)
}
```

Also not bad, misses a few but not much

And if we use fewer seasonal dummies:

```{r plot_bfast_trendseason_02_bic}
for (Sample in ChangeSamples[seq(1, length(ChangeSamples), 10)])
{
    print(Sample)
    SampleTS = GetMatrixFromSF(TreesChange[TreesChange$sample_id==Sample,])[1,]
    MODDetectBreaks(GetTS(SampleTS, frequency = 23), order=1, scrange=NULL, plot=TRUE, formula=response~trend+season, seasonfreq = 0.2, breaks = "BIC", main=Sample, TargetYears=ROWData[ROWData$sample_id==Sample & ROWData$change_at_300m=="yes", ]$reference_year)
}
```

Not as good, too many breaks.

# Mass statistics

Let's run bfast0n using the above and see what the statistics are:

```{r baseline-stat-comparison, cache=TRUE}
Result = TestParams(list(
    list(order=3, scrange=NULL),
    list(order=1, formula=response~trend, scrange=NULL),
    list(breaks="BIC", order=2),
    list(breaks="BIC", order=1, seasonfreq = 0.3, formula=response~trend+season, scrange=NULL)
    ), data=ROWData)
```

```{r baseline-stats, results = "hide", cache=TRUE}
ResultBICO2 = TestMODBreakpointDetection(ROWData, breaks="BIC", order=2, scrange=NULL)
```

```{r}
datatable(rbind(FPStatsPerParam(Result), `breaks BIC, order 2, scrange NULL`=FPStats(ResultBICO2)))
datatable(FPStatsPerClass(ResultBICO2))
```

It appears that seasonal dummies are overall better, check bins

```{r seasonal-comparison, cache=TRUE}
ResultSeason = TestParams(list(
    list(breaks="LWZ", order=1, seasonfreq = 0.3, formula=response~trend+season, scrange=NULL),
    list(breaks="BIC", order=1, seasonfreq = 0.3, formula=response~trend+season),
    list(breaks="BIC", order=1, seasonfreq = 0.2, formula=response~trend+season, scrange=NULL),
    list(breaks="BIC", order=1, seasonfreq = 0.5, formula=response~trend+season, scrange=NULL)
    ), data=ROWData)

```

```{r}
datatable(FPStatsPerParam(ResultSeason))
```


It seems that BIC with 7 bins (0.3*23) works best if we want to keep a balance between sensitivity and specificity. How does that affect the per-class statistics?

```{r perclass-stats}
for (call in unique(Result$call))
{
    print(call)
    datatable(FPStatsPerClass(Result[Result$call==call,]))
}
```
